Illuminating a Dark Lens : A Type Ia Supernova \\ Magnified by Galaxy Cluster Abell 2744

### 2015-04-30 TABLE/magnifications.tex
included tension vs N fig, rearranging sec 6

\begin{deluxetable*}{p{0.7in}cccp{1.4in}p{2.5in}}
\tablecolumns{6}

@@

\tablehead{ \colhead{Model} & \colhead{Best\tablenotemark{a}} & \colhead{Median\tablenotemark{b}} & \colhead{68\% Conf. Range\tablenotemark{c}} & \colhead{References} & \colhead{Description}}

@@

Sharon(v1)   & 2.56     & 2.60  &   2.44$-$2.78 &   HFF\tablenotemark{d}  & LENSTOOL\tablenotemark{e} parametric, strong-lensing based model\\
Sharon(v2)   & 2.74     & 2.59  &   2.42$-$2.85 &   HFF\tablenotemark{d}; \citealt{Johnson:2014} & LENSTOOL parametric, strong-lensing based model. Includes cosmological parameter variations in uncertainty estimates.\\
CATS(v1)     & 2.28     & 2.29  &   2.25$-$2.34 &   HFF\tablenotemark{d}; \citealt{Richard:2014} &  CATS\tablenotemark{f} team implementation of LENSTOOL parametric strong-lensing based model.\\
CATS(v1sw)   & \nodata  & 2.62  &   2.44$-$2.80 &   HFF\tablenotemark{d}; \citealt{Richard:2014} &  CATS\tablenotemark{f} team LENSTOOL parametric model with both strong and weak lensing constraints.\\
CATS(v2)     & \nodata  & 3.42  &   3.27$-$3.58 &   HFF\tablenotemark{d}; \citealt{Jauzac:2014} & Updated version of the CATSv1 model, adds 33 new multiply-imaged galaxies, for a total of 159 individual lensed images.\\
GLAFIC       & 2.34     & 2.29  &   2.19$-$2.37 &   HFF\tablenotemark{d}; \citealt{Ishigaki:2015} & Parametric strong-lensing model using v1.0 of the {\tt GLAFIC} code.\tablenotemark{g} \\
Zitrin-NFW   & 2.09     & 2.29  &   2.07$-$2.52 &   HFF\tablenotemark{d}; \citealt{Zitrin:2013a} &  Parametric strong-lensing model using PIEMD\tablenotemark{h} profiles for galaxies and NFW\tablenotemark{i} profiles for dark matter halos.\\
Zitrin-LTM   & 2.67     & 2.99  &   2.61$-$3.77 &   HFF\tablenotemark{d}; \citealt{Zitrin:2009a} & Parametric strong-lensing model, adopts the Light-Traces-Mass assumption for both the luminous and dark matter.\\
Bradac(v1)   & 3.19     & 2.48  &   2.31$-$2.66 &   HFF\tablenotemark{d}; \citealt{Bradac:2009} & {\tt SWUnited}\tablenotemark{j} : Free-form, strong+weak-lensing based model. Errors from bootstrap resampling only weak-lensing constraints.\\
Bradac(v2)   & 2.23     & 2.26  &   2.30$-$2.23 &   \citealt{Wang:2015} & Updated version of the {\tt SWUnited}\tablenotemark{j} model with new strong-lensing constraints from HFF imaging. Errors from bootstrap resampling only strong-lensing constraints.\\
Williams     & 2.70     & 2.81  &   1.65$-$5.54 &   HFF\tablenotemark{d} & {\tt GRALE}\tablenotemark{k} : Free-form strong-lensing model using a genetic algorithm.  \\
Merten       & 2.33     & 2.24  &   2.04$-$2.92 &   HFF\tablenotemark{d}; \citealt{Merten:2011} &  {\tt SaWLENS},\tablenotemark{l} Grid-based free-form strong+weak lensing based model using adaptive mesh refinement.\\
Lam          & \nodata  & 2.79  &   2.42$-$3.16 &   \citealt{Lam:2014} & {\tt WSLAP+}\tablenotemark{m} : Free-form strong-lensing model using a grid-based method, supplemented by deflections fixed to cluster member galaxies.\\
Diego\tablenotemark{n}  & \nodata  & 1.80  &   1.44$-$2.16 & \citealt{Diego:2014b} & Alternative implementation of the {\tt WSLAP+}\tablenotemark{m} model, using a different set of strong-lensing constraints and redshifts.

@@

\tablenotetext{a}{The magnification returned for the optimal version of each model, as independently defined by each lens modeling team.}  
\tablenotetext{b}{Median magnification from 100-600 Monte Carlo realizations of the model.}  
\tablenotetext{c}{Confidence ranges about the median, enclosing 68\%\ of the realized values.}
\tablenotetext{d}{Lens models with the reference code ``HFF'' were produced as part of the 

@@

\tablenotetext{e}{{\tt LENSTOOL} : \citealt{Jullo:2007}; \url{http://projects.lam.fr/repos/lenstool/wiki}}
\tablenotetext{f}{{\tt CATS} : Clusters As TelescopeS lens modeling team. PI's: J.-P. Kneib \&\ P. Natarajan}
\tablenotetext{g}{{\tt GLAFIC} : \citealt{Oguri:2010}; \url{http://www.slac.stanford.edu/~oguri/glafic/}}
\tablenotetext{h}{PIEMD: Pseudo Isothermal Elliptical Mass Distrubition}
\tablenotetext{i}{NFW : Navarro-Frenk-White mass density profile \citep{Navarro:1997}.}
\tablenotetext{j}{{\tt SWunited} : Strong and Weak lensing United; \citealt{Bradac:2005}}
\tablenotetext{k}{{\tt GRALE} : GRAvitational LEnsing; \citealt{Liesenborgs:2006,Liesenborgs:2007,Mohammed:2014}}
\tablenotetext{l}{{\tt SaWLENS} : \citealt{Merten:2009}; Strong and Weak LENSing analysis code. \url{http://www.julianmerten.net/codes.html}}
\tablenotetext{m}{{\tt WSLAP+} : \citealt{Sendra:2014}; Weak and Strong Lensing Analysis Package plus member galaxies (Note: no weak-lensing constraints used for Abell 2744)}
\tablenotetext{n}{Abell 2744 model available at \url{http://www.ifca.unican.es/users/jdiego/LensExplorer}. 

@@

>>>>

\clearpage
\begin{turnpage}
\begin{deluxetable*}{p{0.62in}rrrrcccp{1.3in}p{2.9in}}
\tablecolumns{10}

@@

\tablehead{ \colhead{Model} & \colhead{Sys\tablenotemark{a}} & \colhead{Im\tablenotemark{a}} & 
            \colhead{Spec\tablenotemark{b}} & \colhead{Phot\tablenotemark{b}} & 
            \colhead{Best\tablenotemark{c}} & \colhead{Median\tablenotemark{d}} & 
            \colhead{68\% Conf. Range\tablenotemark{e}} & \colhead{References} & \colhead{Description}}

@@

Sharon(v1)     & 17 &  60 & 2 & 14 & 2.56     & 2.60  &   2.44$-$2.78 &   HFF\tablenotemark{f}  & LENSTOOL\tablenotemark{g} parametric, strong-lensing based model\\
Sharon(v2)     & 17 &  60 & 2 & 14 & 2.74     & 2.59  &   2.42$-$2.85 &   HFF; \citealt{Johnson:2014} & LENSTOOL parametric, strong-lensing based model. Includes cosmological parameter variations in uncertainty estimates.\\
CATS(v1)       & 50 & 151 & 4 & 0  & 2.28     & 2.29  &   2.25$-$2.34 &   HFF; \citealt{Richard:2014} &  CATS\tablenotemark{h} team implementation of LENSTOOL parametric strong-lensing based model.\\
CATS(v1sw)     & 24 &  67 & 3 & 12 & \nodata  & 2.62  &   2.44$-$2.80 &   HFF; \citealt{Richard:2014} &  CATS team LENSTOOL parametric model with both strong and weak lensing constraints.\\
CATS(v2)       & 17 &  60 & 2 & 14 & \nodata  & 3.42  &   3.27$-$3.58 &   HFF; \citealt{Jauzac:2014c} & Updated version of the CATSv1 model, adds 33 new multiply-imaged galaxies, for a total of 159 individual lensed images.\\
GLAFIC         & 15 &  47 & 3 & 11 & 2.34     & 2.29  &   2.19$-$2.37 &   HFF; \citealt{Ishigaki:2015} & Parametric strong-lensing model using v1.0 of the {\tt GLAFIC} code.\tablenotemark{i} \\
Zitrin-NFW     & 10 &  44 & 2 & 0  & 2.09     & 2.29  &   2.07$-$2.52 &   HFF; \citealt{Zitrin:2013a} &  Parametric strong-lensing model using PIEMD\tablenotemark{j} profiles for galaxies and NFW\tablenotemark{k} profiles for dark matter halos.\\
Zitrin-LTM     & 10 &  44 & 2 & 0  & 2.67     & 2.99  &   2.61$-$3.77 &   HFF; \citealt{Zitrin:2009a} & Parametric strong-lensing model, adopts the Light-Traces-Mass assumption for both the luminous and dark matter.\\
Bradac(v1)     & 16 &  56 & 2 & 14 & 3.19     & 2.48  &   2.31$-$2.66 &   HFF; \citealt{Bradac:2009} & {\tt SWUnited}\tablenotemark{l} : Free-form, strong+weak-lensing based model. Errors from bootstrap resampling only weak-lensing constraints.\\
Bradac(v2)     & 10 &  40 & 2 & 8  & 2.23     & 2.26  &   2.30$-$2.23 &   \citealt{Wang:2015} & Updated version of the {\tt SWUnited} model with new strong-lensing constraints from HFF imaging. Errors from bootstrap resampling only strong-lensing constraints.\\
Williams       & 16 &  56 & 2 & 14 & 2.70     & 2.81  &   1.65$-$5.54 &   HFF & {\tt GRALE}\tablenotemark{m} : Free-form strong-lensing model using a genetic algorithm.  \\
Merten         & 25 &  72 & 7 & 18 & 2.33     & 2.24  &   2.04$-$2.92 &   HFF; \citealt{Merten:2011} &  {\tt SaWLENS},\tablenotemark{n} Grid-based free-form strong+weak lensing based model using adaptive mesh refinement.\\
Lam(v1)        & 21 &  65 & 4 & 17 & \nodata  & 2.79  &   2.42$-$3.16 &   \citealt{Lam:2014} & {\tt WSLAP+}\tablenotemark{o} : Free-form strong-lensing model using a grid-based method, supplemented by deflections fixed to cluster member galaxies.\\
Lam(v2)        & 10 &  32 & 5 & 5  & 1.86     & 1.91  &   1.54$-$2.28 &   \citealt{Lam:2014} & Updated version of the {\tt WSLAP+} model, using more selective strong-lensing constraints and GALFIT\tablenotemark{p} models for galaxy mass.\\
Diego$^{\rm q}$ & 15 &  48 & 4 & 11 & \nodata  & 1.80  &   1.44$-$2.16 & \citealt{Diego:2014b} & Alternative implementation of the {\tt WSLAP+} model, using a different set of strong-lensing constraints and redshifts.

@@

\tablenotetext{a}{Number of multiply imaged systems used as strong-lensing constraints.}
\tablenotetext{b}{Number of multiply imaged systems with spectroscopic redshifts.}
\tablenotetext{c}{The magnification returned for the optimal version of each model, as independently defined by each lens modeling team.}  
\tablenotetext{d}{Median magnification from 100-600 Monte Carlo realizations of the model.}  
\tablenotetext{e}{Confidence ranges about the median, enclosing 68\%\ of the realized values.}
\tablenotetext{f}{Lens models with the reference code ``HFF'' were produced as part of the 

@@

\tablenotetext{g}{{\tt LENSTOOL} : \citealt{Jullo:2007}; \url{http://projects.lam.fr/repos/lenstool/wiki}}
\tablenotetext{h}{{\tt CATS} : Clusters As TelescopeS lens modeling team. PI's: J.-P. Kneib \&\ P. Natarajan}
\tablenotetext{i}{{\tt GLAFIC} : \citealt{Oguri:2010}; \url{http://www.slac.stanford.edu/~oguri/glafic/}}
\tablenotetext{j}{PIEMD: Pseudo Isothermal Elliptical Mass Distrubition}
\tablenotetext{k}{NFW : Navarro-Frenk-White mass density profile \citep{Navarro:1997}.}
\tablenotetext{l}{{\tt SWunited} : Strong and Weak lensing United; \citealt{Bradac:2005}}
\tablenotetext{m}{{\tt GRALE} : GRAvitational LEnsing; \citealt{Liesenborgs:2006,Liesenborgs:2007,Mohammed:2014}}
\tablenotetext{n}{{\tt SaWLENS} : \citealt{Merten:2009}; Strong and Weak LENSing analysis code. \url{http://www.julianmerten.net/codes.html}}
\tablenotetext{o}{{\tt WSLAP+} : \citealt{Sendra:2014}; Weak and Strong Lensing Analysis Package plus member galaxies (Note: no weak-lensing constraints used for Abell 2744)}
\tablenotetext{p}{{\tt GALFIT} : Two-dimensional galaxy fitting algorithm \citep{Peng:2002}}
\tablenotetext{q}{Abell 2744 model available at \url{http://www.ifca.unican.es/users/jdiego/LensExplorer}. 

@@

\end{turnpage}
\clearpage
\global\pdfpageattr\expandafter{\the\pdfpageattr/Rotate 90}

### 2015-04-30 snTomas.tex
included tension vs N fig, rearranging sec 6

\begin{abstract}
SN \tomas\ is a Type Ia Supernova (\SNIa) discovered at
\change{$z=1.3457\pm0.0001$} behind the galaxy cluster Abell 2744 ($z=0.308$). This
SN has a projected separation from the cluster core of
$\sim$40\arcsec, closer to the critical lensing line than any previous
cluster-lensed Type Ia SN.  As such, it provides a valuable
opportunity to confront gravitational lens models of galaxy clusters
with the direct measurement of an absolute magnification at the edge
of the strong-lensing region.  We derive a tightly constrained measure
of the peak apparent magnitude, corrected for light curve shape and
extinction.  In a cosmology-independent analysis, we find that \tomas\
is \change{$0.77\pm0.15$} magnitudes brighter than unlensed Type Ia
SNe at similar redshift, implying a lensing magnification
of \change{$\mu_{\rm obs}=2.03\pm0.29$}.  We compare this direct
measurement of the magnification against predicted magnifications
from 14 lens models, representing a broad range of
methodologies and incorporating a variety of strong- and weak-lensing
constraints.  The models are collectively fairly accurate: 6 of the
models deliver median magnifications that are consistent with the
measured $\mu$ to within the 1$\sigma$ uncertainties.  However, the
models are systematically biased to higher values: \change{all but
one} of the tested models yield a median magnification higher than the
observed $\mu$, and \change{6 of the models are $\geq1.5\sigma$}
above the measured magnification.  We evaluate possible causes for
this mild bias, considering both modeling errors and
misinterpretation of the SN data. Although we find no single
explanation that can resolve this, we \change{ highlight several ways
that lens models can deliver more realistic uncertainties, such as
incorporating errors from assumed cosmological parameters and allowing
greater flexibility in the form of the cluster mass distribution.  A
more robust evaluation of systematic biases could be achieved with
further cooperative lens model comparisons using a larger sample of
lensed Type Ia SNe and cluster lensing simulations.}
\end{abstract}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{FIG/snTomas_hubble_diagram}
\caption{ Measurement of the lensing magnification from comparison of
the \tomas\ distance modulus to a sample of unlensed field SN.  SNe
from the GOODS program are plotted as squares, those from the SCP
survey are shown as circles, and the CANDELS objects are triangles.
The distance modulus for each SN is derived from light curve fits
using the MLCS2k2 fitter (top panel) and the SALT2 fitter (bottom
panel).
\label{fig:MagnificationMeasurement} }
\end{center}
\end{figure}

@@

Before the Frontier Fields observations began, the Space Telescope
Science Institute (STScI) issued a call for lens modeling teams to
generate mass models of all 6 Frontier Field clusters, using a shared
collection of all imaging and spectroscopic data available at the
time.  In response to this opportunity, five teams generated eight
models for Abell 2744.  These models necessarily relied on
pre-HFF data, and were required to be complete before the HFF program
began, in order to enable the estimation of magnifications for any new
lensed background sources revealed by the HFF imaging. An interactive
web tool was created by D. Coe and hosted at STScI, to extract
magnification estimates and uncertainties from each model for any
given redshift and position.  In this work we also consider three
additional models that were created later, taking advantage of new
multiply-imaged galaxies discovered in the HFF imaging as well as new
redshifts for lensed background galaxies.

\input{TABLE/magnifications}

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{FIG/snTomas_magnifications}
\caption{ 
Comparison of the observed lensing magnification to predictions from
lens models. The vertical blue line shows the constraints from
SN \tomas\ derived in Section~\ref{sec:DistanceAndMagnification} using the
MLCS2k2 fitter, with a shaded region marking the total
uncertainty. Markers with horizontal error bars show the median
magnification and 68\%\ confidence region from each of the 14 lensing
models.  Circles indicate models that use only strong-lensing
constraints, while diamonds denote those that also incorporate
weak-lensing measurements.  Models shown with open markers were
constructed using only data available before the start of the Frontier
Fields observations.  Filled markers indicate models that use
additional input constraints, including new multiply imaged systems
and redshifts.  The eight models shown in black at the top are from
the ``parametric'' family, and the six in green at the bottom are
``free-form.''  The grey dashed line marks the unweighted mean for the
set of all models, at $\mu=2.53$.
\label{fig:lensingtest} }
\end{center}
\end{figure}

Table~\ref{tab:PredictedMagnifications} lists the 14 models,
giving for each one the predicted magnification and uncertainty for a
source at \change{$z=1.3457$} and at the position of \tomas.  These models
represent a broad sampling of the techniques and assumptions that can
be applied to the modeling of mass distributions in galaxy clusters. A
rigorous comparison of these disparate lens modeling techniques is
beyond the scope of this work. Table~\ref{tab:PredictedMagnifications}
includes a brief description of each model, but for a complete
discussion of the methodology the reader is referred to the listed
references.

In Figure~\ref{fig:lensingtest} the model predictions are plotted
alongside the observed magnification of SN \tomas, derived in
Section~\ref{sec:DistanceAndMagnification}.  This comparison shows that these 14
models are largely consistent with each other. Naively treating each
model as an independent prediction for the magnification (and ignoring
their quoted uncertainties), one would get a mean for the full sample
of \change{$\mu=2.53\pm0.38$.  This is separated from the observed SN
magnification by $\delta\mu/\mu=25\%$, which is just over a
1$\sigma$ difference.}

The models shown in Figure~\ref{fig:lensingtest} are separated along
the $y$ direction into two broad classes, with eight so-called {\it
parametric}\footnote{Although this nomenclature is becoming standard
in the literature, it is somewhat misleading, as the pixels or grid
cells in free-form models are effectively parameters as well. Perhaps
``simply-parameterized'' would be more accurate, though we adopt the
more common usage here.}  models on top and six {\it free-form} models
on the bottom.  Broadly speaking, the parametric models use
parameterized density distributions to describe the arrangement of
mass within the cluster, typically associating massive dark matter
halos with the positions of bright cluster member galaxies. Therefore,
parametric models rely (to varying degrees) on the assumption that the
cluster's dark matter can be described by analytic forms such as NFW
halos \citep{Navarro:1997}, or pseudo isothermal elliptical mass
distributions \citep{Kassiola:1993}.  Free-form models divide the
cluster into a grid, generally using a
multi-scale grid to get better sampling in denser regions. Each grid cell
is assigned a mass or a potential, and then the mass values are
iteratively refined to match the observed lensing
constraints. In some cases an adaptive grid is used so
that the grid spacing itself can also be modified as the model is
iterated \citep[e.g.][]{Liesenborgs:2006,Merten:2009}.

In general, free-form methods tend to explore the model parameter
space more fully than parametric methods.  This is reflected in the
size of the error-bars in Figure~\ref{fig:lensingtest}.  For example,
the Williams model (using GRALE), which has the largest, and possibly the
most realistic uncertainties is able to account for mass sheet
degeneracy by introducing a free parameter in the reconstruction that
represents an arbitrarily scaled mass sheet, with normalization to
be chosen by the genetic algorithm.  More complex degeneracies are
also accounted for in this model through the use of basis
functions \citep{Liesenborgs:2006,Liesenborgs:2007,Mohammed:2014}

Figure~\ref{fig:lensingtest} also separates the models based on the
scope of input data constraints used.  Ten of the models rely only on
strong-lensing constraints (plotted as circles), while the other four
also use weak-lensing measurements (diamonds).  Nine of the models
were constructed using only pre-HFF data (open symbols), while five
others added new multiply-imaged systems and redshifts.

In spite of the great variety in input data and methodology, these
lens models overall are delivering consistent and fairly accurate
estimates of the magnification.  Six of the tested models
(CATSv1, GLAFIC, Zitrin-NFW, Bradac(v2), Merten, Diego) report a median
value that falls within the 1$\sigma$ uncertainty range of the
measured magnification, and two of the free-form models (Williams,
Diego) report uncertainties that overlap the central $\mu$ as measured
from both of the SN light curve fitters.  Furthermore, the scatter
amongst models provides a reasonable estimate of the magnification
uncertainty.

This is largely in agreement with the results
of \citetalias{Patel:2014} and \citet{Nordin:2014}.  The fact that the
model predictions are collectively within $2\sigma$ of the observed
magnification is especially encouraging for these Abell 2744
models. This is a merging cluster with a complex mass distribution,
and most of the models we are evaluating are preliminary products,
generated quickly and without access to the rich data from the
Frontier Fields program.

However, beyond this first-order agreement, there is a small
systematic bias apparent. All of the lens models \change{except one}
return median magnifications that are {\it higher} than the observed
value, and \change{six of the models are discrepant by more than $1.5\sigma$.}
It is important to emphasize that SN \tomas\ only samples a single
sight-line through the cluster, so any conclusions to be drawn from
this analysis are necessarily limited.  Nevertheless, a systematic
shift common to all models is surprising, given the wide range of
modeling strategies, input data, and physical assumptions represented
by this set of models.  In the following subsections we examine
possible explanations for this small but \change{nearly} universal
bias.

@@

\subsection{Redshift Error}
\label{sec:RedshiftError}

If the redshift of the SN derived in Section~\ref{sec:Spectroscopy}
were incorrect, then one would derive a different value for the
magnification, both from the SN measurement and the lens model
predictions.  Conceivably, this could resolve the tension between the
measurement and the models.  It is often the case in SN surveys that
redshifts are assigned based on a host galaxy association, typically
inferred from the projected separation between the SN and nearby
galaxies.  In this case the redshift evidence \change{is strongly supported} with evidence from the SN
itself: we find a consistent redshift from both the SN spectrum
(Section~\ref{sec:SpecRedshift}) and the light curve
Section~\ref{sec:PhotometricClassification}, which are both within the
$1\sigma$ range of the photometric redshift for the nearest detected
galaxy: $z=1.5\pm0.2$. This appears to be a solid and self-consistent
picture, so the evidence strongly disfavors any redshift that is
significantly different from $z=1.35$.

\subsection{Foreground Dust and SN Color}
\label{sec:ForegroundDust}

\subsection{Misclassification}
\label{sec:Misclassification}

@@

\end{document}

>>>>

\begin{abstract}
SN \tomas\ is a Type Ia Supernova (\SNIa) discovered at
\change{$z=1.3457\pm0.0001$} behind the galaxy cluster Abell 2744 ($z=0.308$). This
SN has a projected separation from the cluster core of
$\sim$40\arcsec, closer to the critical lensing line than any previous
cluster-lensed Type Ia SN.  As such, it provides a valuable
opportunity to confront gravitational lens models of galaxy clusters
with the direct measurement of an absolute magnification at the edge
of the strong-lensing region.  We derive a tightly constrained measure
of the peak apparent magnitude, corrected for light curve shape and
extinction.  In a cosmology-independent analysis, we find that \tomas\
is \change{$0.77\pm0.15$} magnitudes brighter than unlensed Type Ia
SNe at similar redshift, implying a lensing magnification
of \change{$\mu_{\rm obs}=2.03\pm0.29$}.  We compare this direct
measurement of the magnification against predicted magnifications
from \changeR{15 lens models developed by 8 independent groups},
representing a broad range of methodologies and incorporating a
variety of strong- and weak-lensing constraints.  The models are
collectively fairly accurate: 8 of the models deliver median
magnifications that are consistent with the measured $\mu$ to within
the 1$\sigma$ uncertainties.  \changeR{However, the models exhibit a
subtle systematic bias: all of the models that disagree with the SN by
more than $1\sigma$ are {\it higher} than the measured value}
and \change{6 of these models are $\geq1.5\sigma$} above the measured
magnification.  We evaluate possible causes for this mild
bias, \changeR{and reject explanations that invoke modeling errors or
misinterpretation of the SN data. We conclude that -- at least for
this single line of sight -- the most accurate models are those with
(1) stringent quality cuts for strong-lensing constraints and (2)
relatively flexible model design}. \change{A more robust evaluation of
systematic biases could be achieved with further cooperative lens
model comparisons using a larger sample of lensed Type Ia SNe as well
as cluster lensing simulations.}
\end{abstract}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{FIG/snTomas_hubble_diagram}
\caption{ Measurement of the lensing magnification from comparison of
the \tomas\ distance modulus to a sample of unlensed field SN.  SNe
from the GOODS program are plotted as squares, those from the SCP
survey are shown as circles, and the CANDELS objects are triangles.
The distance modulus for each SN is derived from light curve fits
using the MLCS2k2 fitter (top panel) and the SALT2 fitter (bottom
panel).  \changeR{Grey lines with shading show a linear fit to the
unlensed control sample is shown, anchored at the redshift of
SN \tomas, with fit parameters given at the top.  The derived
magnification is reported at the bottom of each panel.}
\label{fig:MagnificationMeasurement} }
\end{center}
\end{figure}

@@

Before the Frontier Fields observations began, the Space Telescope
Science Institute (STScI) issued a call for lens modeling teams to
generate mass models of all 6 Frontier Field clusters, using a shared
collection of all imaging and spectroscopic data available at the
time.  In response to this opportunity, five teams generated seven
models for Abell 2744.  These models necessarily relied on
pre-HFF data, and were required to be complete before the HFF program
began, in order to enable the estimation of magnifications for any new
lensed background sources revealed by the HFF imaging. An interactive
web tool was created by D. Coe and hosted at STScI, to extract
magnification estimates and uncertainties from each model for any
given redshift and position.  In this work we also consider 8
additional models that were created later.  \changeR{Some of these are
updates of the original models produced in response to the lens
modeling call,  and some of them include new
multiply-imaged galaxies discovered in the HFF imaging as well as new
redshifts for lensed background galaxies.  Magnifications derived from
several of these later models are also available on the interactive
web tool hosted by STScI.  The full list of models and details on
their construction are given in Table~\ref{tab:LensModels}.}

\input{TABLE/magnifications2}

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{FIG/snTomas_magnifications}
\caption{ 
Comparison of the observed lensing magnification to predictions from
lens models. The vertical blue line shows the constraints from
SN \tomas\ derived in Section~\ref{sec:DistanceAndMagnification} using the
MLCS2k2 fitter, with a shaded region marking the total
uncertainty. Markers with horizontal error bars show the median
magnification and 68\%\ confidence region from each of the 15 lensing
models.  Circles indicate models that use only strong-lensing
constraints, while diamonds denote those that also incorporate
weak-lensing measurements.  Models shown with open markers were
constructed using only data available before the start of the Frontier
Fields observations.  Filled markers indicate models that use
additional input constraints, including new multiply imaged systems
and redshifts.  Points with magenta outline indicate ``unblind''
models, which were generated after the magnification of the SN was
known. The eight models shown in black at the top are from
the ``parametric'' family, and the six in green at the bottom are
``free-form.''  The grey dashed line marks the unweighted mean for the
set of all models, at $\mu=2.5$.
\label{fig:lensingtest} }
\end{center}
\end{figure}

\changeR{
Table~\ref{tab:PredictedMagnifications} gives each model's predicted}
magnification and uncertainty for a source at \change{$z=1.3457$} and
at the position of \tomas.  These models represent a broad sampling of
the techniques and assumptions that can be applied to the modeling of
mass distributions in galaxy clusters. A rigorous comparison of these
disparate lens modeling techniques is beyond the scope of this
work. Table~\ref{tab:PredictedMagnifications} includes a brief
description of each model, but for a complete discussion of the
methodology the reader is referred to the listed references.

In Figure~\ref{fig:lensingtest} the model predictions are plotted
alongside the observed magnification of SN \tomas, derived in
Section~\ref{sec:DistanceAndMagnification}.  This comparison shows that these 14
models are largely consistent with each other. Naively treating each
model as an independent prediction for the magnification (and ignoring
their quoted uncertainties), one would get a mean for the full sample
of \change{$\mu=2.5\pm0.4$.  This is separated from the observed SN
magnification by $\delta\mu/\mu=23\%$, which is approximately a
1$\sigma$ difference.}

The models shown in Figure~\ref{fig:lensingtest} are separated along
the $y$ direction into two broad classes, with eight so-called {\it
parametric}\footnote{Although this nomenclature is becoming standard
in the literature, it is somewhat misleading, as the pixels or grid
cells in free-form models are effectively parameters as well. Perhaps
``simply-parameterized'' would be more accurate, though we adopt the
more common usage here.}  models on top and seven {\it free-form} models
on the bottom.  Broadly speaking, the parametric models use
parameterized density distributions to describe the arrangement of
mass within the cluster, typically associating massive dark matter
halos with the positions of bright cluster member galaxies. Therefore,
parametric models rely (to varying degrees) on the assumption that the
cluster's dark matter can be described by analytic forms such as NFW
halos \citep{Navarro:1997}, or pseudo isothermal elliptical mass
distributions \citep{Kassiola:1993}.  Free-form models divide the
cluster into a grid, generally using a
multi-scale grid to get better sampling in denser regions. Each grid cell
is assigned a mass or a potential, and then the mass values are
iteratively refined to match the observed lensing
constraints. In some cases an adaptive grid is used so
that the grid spacing itself can also be modified as the model is
iterated \citep[e.g.][]{Liesenborgs:2006,Merten:2009}.

In general, free-form methods tend to explore the model parameter
space more fully than parametric methods.  This is reflected in the
size of the error-bars in Figure~\ref{fig:lensingtest}.  For example,
the Williams model (using GRALE), which has the largest uncertainties
is able to account for mass sheet degeneracy by introducing a free
parameter in the reconstruction that represents an arbitrarily scaled
mass sheet, with normalization to be chosen by the genetic algorithm.
More complex degeneracies are also accounted for in this model through
the use of basis
functions \citep{Liesenborgs:2006,Liesenborgs:2007,Mohammed:2014}

Figure~\ref{fig:lensingtest} also separates the models based on the
scope of input data constraints used.  Ten of the models rely only on
strong-lensing constraints (plotted as circles), while the other four
also use weak-lensing measurements (diamonds).  Nine of the models
were constructed using only pre-HFF data (open symbols), while six
others added new multiply-imaged systems and redshifts (filled symbols).
For twelve of these models the SN magnification qualifies as a true
``blind test'', as they were completed before the SN \tomas\
magnification measurement was known.  The remaining three models
(Bradac-v2, Lam-v2, and Diego) are technically not blind, although
none of these modelers used the SN magnification as a constraint, and
the modelers did not have access to the final SN magnification value
when constructing their model.

In spite of the great variety in input data and methodology, these
lens models overall are delivering consistent and fairly accurate
estimates of the magnification.  Seven of the tested models (CATS(v1),
GLAFIC, Zitrin-NFW, Merten, Bradac(v2), Lam(v2), Diego) report a
median value that falls within the 1$\sigma$ uncertainty range of the
measured magnification, and three of the free-form models (Williams,
Lam(v2), Diego) report uncertainties that overlap the central $\mu$ as
measured from both of the SN light curve fitters.  Furthermore, the
scatter amongst models provides a reasonable estimate of the
magnification uncertainty.

This is largely in agreement with the results
of \citetalias{Patel:2014} and \citet{Nordin:2014}.  The fact that the
model predictions are collectively within $1\sigma$ of the observed
magnification is especially encouraging for these Abell 2744
models. This is a merging cluster with a complex mass distribution,
and many of the models we are evaluating are preliminary products,
generated quickly and without access to the rich data from the
Frontier Fields program.

However, beyond this first-order agreement, there is a small
systematic bias apparent. All but two of the lens models
return median magnifications that are {\it higher} than the observed
value, and \change{six of the models are discrepant by more than $1.5\sigma$.}
It is important to emphasize that SN \tomas\ only samples a single
line of sight through the cluster, and this bias to higher
magnifications is minor.  Nevertheless, a systematic
shift of this nature is surprising, given the wide range of
modeling strategies, input data, and physical assumptions represented
by this set of models.  In the following subsections we examine
possible explanations for this small but \change{nearly} universal
bias.

\subsection{Possible Errors in Supernova Interpretation}
\label{sec:SupernovaErrors}

\changeR{Before examining the lens models or the cluster, let us first
consider whether we might have misinterpreted the data on the SN
itself.}

\subsubsection{Redshift Error}
\label{sec:RedshiftError}

If the redshift of the SN derived in Section~\ref{sec:Spectroscopy}
were incorrect, then one would derive a different value for the
magnification, both from the SN measurement and the lens model
predictions.  Conceivably, this could resolve the tension between the
measurement and the models.  It is often the case in SN surveys that
redshifts are assigned based on a host galaxy association, typically
inferred from the projected separation between the SN and nearby
galaxies.  In this case the redshift \change{is strongly supported} by
evidence from the SN
itself: we find a consistent redshift from both the SN spectrum
(Section~\ref{sec:SpecRedshift}) and the light curve
Section~\ref{sec:PhotometricClassification}, which are both within the
$1\sigma$ range of the photometric redshift for the nearest detected
galaxy: $z=1.5\pm0.2$. This appears to be a solid and self-consistent
picture, so the evidence strongly disfavors any redshift that is
significantly different from $z=1.35$.

\subsubsection{Foreground Dust and SN Color}
\label{sec:ForegroundDust}

\subsubsection{Misclassification}
\label{sec:Misclassification}

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{FIG/tension}
\caption{ \label{fig:Nspecz}
\changeR{
The model tension as a function of strong-lensing constraints. The y
axis in both panels marks the number of standard deviations in
discrepancy between the measured SN magnification and the prediction
from each lens model. The left panel plots this tension against the
total number of multiply imaged systems used, and the right panel
plots the fraction of
multiply-imaged systems that have a spectroscopic redshift
constraint.  Marker shapes and colors are the same as in
Figure~\ref{fig:LensingTest}.}
}
\end{center}
\end{figure*}

\subsection{Possible Error in Lens Model Construction}
\label{sec:LensModelErrors}

\changeR{Having now ruled out misinterpetations of the SN
data, we are left to seek an explanation for the small systematic bias
by scrutinizing the lens models.  TODO: note that 
the most successful models are those that use a
stringent criteria for strong lensing constraints, which results in a
larger fraction that have spectroscopic redshifts. 
}

@@

@@

%\input{TABLE/magnifications}
\end{document}
